\chapter{Zero free region for the zeta function}\label{zerofree-chapter}

\unintegrated 

A zero of the Riemann zeta function is a complex number $\rho = \beta + i\gamma$ for which $\zeta(\rho) = 0$. The zeta function has a infinite number of zeros of the form $\rho = -2n$ for integer $n \ge 1$; these are known as trivial zeros and are well understood. There are also an infinite number zeros inside the ``critical strip" $0 < \Re z < 1$, called non-trivial zeros. The locations of the non-trivial zeros have deep consequences for many fields of mathematics. A well-known conjecture regarding the non-trivial zeros is the Riemann hypothesis. 

\begin{conjecture}[Riemann hypothesis]\label{rh}
If $\rho$ is a non-trivial zero of the Riemann zeta function, then $\Re \rho = 1/2$.
\end{conjecture}

This conjecture remains far out of reach. Instead, currently what are known are zero free regions.  

\begin{definition}[Zero free region]\label{zeta-zero-free-def}
A zero-free region of the Riemann zeta function is a set $D \subset \mathbb{C}$ for which $\zeta(s) \ne 0$ for all $s \in D$. 
\end{definition}

\begin{lemma}[Basic properties of zero free regions]\label{zero-free-basic-lem}
The following properties hold:
\begin{enumerate}
    \item[(i)] (Symmetry about the real axis) If $\zeta(\sigma + it) \ne 0$ then $\zeta(\sigma - it) \ne 0$.
    \item[(ii)] (Symmetry about the critical line $\Re s = 1/2$) For $0 \le \sigma \le 1$, if $\zeta(\sigma + it) \ne 0$ then $\zeta(1 - \sigma + it) \ne 0$.
    \item[(iii)] (Non vanishing for $\Re s > 1$) If $\Re s > 1$ then $\zeta(s) \ne 0$.
\end{enumerate}
\end{lemma}
\begin{proof}
Claim (i) follows directly from the property $\overline{\zeta(s)} = \zeta(\overline{s})$. Claim (ii) follows from the functional equation 
\[
\zeta(s) = 2^s \pi^{s - 1}\sin(\pi s/2) \Gamma(1 -s)\zeta(1-s)
\]
and claim (iii) follows from the Euler product formula
\[
\zeta(s) = \prod_{p}(1 - p^{-s})^{-1},\qquad (\Re s > 1).
\]
\end{proof}

In light of \Cref{zero-free-basic-lem}, for the rest of the chapter we will focus on the quadrant 
\[
D\subseteq \{z \in \mathbb{C}: \Re z > 1/2, \Im z > 0\}.
\]
The first non-trivial zero-free region was due to de la Vall\'ee Poussin \cite{de_la_vallee_poussin_recherches_1896} and Hadamard \cite{hadamard_distribution_1896}, who showed independently that:


\begin{theorem}[Non-vanishing on the 1-line]
One has $\zeta(1 + it) \ne 0$ for any real $t$.
\end{theorem}
\begin{proof}
For $\Re s > 1$, one has
\[
\Re \log \zeta(s) = \sum_{p}\sum_{m = 1}^{\infty}\frac{\cos (t \log p^m)}{mp^{m\sigma}}
\]
where the outer sum runs through all primes. Applying this formula at $s = \sigma, \sigma + it$ and $\sigma + 2it$ ($t \ne 0$), and since $3 + 4\cos\theta + \cos 2\theta = 2(1 + \cos\theta)^2 \ge 0$, one has
\[
3\Re \log \zeta(\sigma) + 4\Re \log \zeta(\sigma + it) + \Re\log \zeta(\sigma + 2it) = \sum_{p}\sum_{m = 1}^{\infty}\frac{3 + \cos (t \log p^m) + \cos (2t\log p^m)}{mp^{m\sigma}} \ge 0.
\]
It follows that $|\zeta(\sigma)^3 \zeta(\sigma + it)^4 \zeta(\sigma + 2it)| \ge 1$. Now as $\sigma \to 1$ from above (and $t$ remains fixed), one has
\[
\zeta(\sigma) \ll \frac{1}{\sigma - 1},\qquad \zeta(\sigma + 2it) \ll 1,
\]
since $\zeta$ has a simple pole at $s = 1$ and no pole at $\sigma + 2it$. If $\zeta(1 + it) = 0$, then $\zeta(\sigma + it) \ll \sigma - 1$ so that $|\zeta(\sigma)^3 \zeta(\sigma + it)^4 \zeta(\sigma + 2it)| \ll \sigma - 1$, a contradiction.
\end{proof}

This was used to prove the prime number theorem $\pi(x) \sim x/\log x$ as $x \to \infty$ (in fact the two statements are equivalent).

\section{Relation to upper bound on zeta in the critical strip}

Using estimates of $\zeta(\sigma + it)$ close to the line $\sigma = 1$, one can extend the zero free region slightly inside the critical strip. 

\begin{lemma}[Relation to growth exponents of zeta]\label{mu_to_zero_free}
Suppose and $0 < g(t) \le 1 < f(t)$ are real-valued functions for $t \ge 0$, with $f(t)$ non-decreasing and tending to infinity with $t$, and $g(t)$ non-increasing. Suppose further that $e^{f(t)}/g(t) = o(f(t))$. If 
\[
\zeta(\sigma + it) \ll f(t)\qquad (1 - g(t) \le \sigma \le 2, t \ge 0)
\]
then $\zeta(\sigma + it) \ne 0$ for 
\[
\sigma \ge 1 - A\frac{g(2t + 1)}{\log f(2t + 1)}
\]
where $A > 0$ is an absolute constant. 
\end{lemma}
\begin{proof}
See \cite[Theorem 3.10]{titchmarsh_theory_1986}.
\end{proof}

\begin{theorem}[Classical zero free region]\label{zfr-classical}
One has $\zeta(\sigma + it) \ne 0$ if 
\[
\sigma \ge 1 - \frac{A}{\log t}.
\]
for an absolute constant $A > 0$ and $t$ sufficiently large.
\end{theorem}
\begin{proof}
Thanks to the convexity bound $\mu(\sigma) \le (1-\sigma)/2$, one may take $g(t) = 1/2$, $f(t) = t^{1/4 + o(1)}$ in \Cref{mu_to_zero_free}. The result follows.
\end{proof}

This classical result has been improved in a number of works, most of which make crucial use of non-trivial estimates of certain types of exponential sums.

\begin{theorem}[Littlewood zero free region]\label{zfr-littlewood}
One has $\zeta(\sigma + it) \ne 0$ if 
\[
\sigma \ge 1 - \frac{A \log\log t}{\log t}
\]
for an absolute constant $A > 0$ and $t$ sufficiently large.
\end{theorem}
\begin{proof}
Follows from the zeta bound corresponding to
\[
\mu\left(1 - \frac{k}{2^k - 2}\right) \le \frac{1}{2^k - 2}
\]
for integer $k\ge 3$, which is generated by the van der Corput exponent pair $A^{k - 2}B(0, 1) = (\frac{1}{2^k - 2}, 1 - \frac{k - 1}{2^k - 2})$. However, one needs to make explicit the $o(1)$ term in the bound $\zeta(\sigma + it) \ll t^{\mu(\sigma) + o(1)}$. In particular, by \cite[Theorem 5.14]{titchmarsh_theory_1986}, one has 
\[
\zeta(1 - \frac{k}{2^k - 2} + it) \ll t^{1/(2^k - 2)}\log t.
\]
Taking 
\[
k = \left\lfloor \frac{1}{\log 2}\log\left(\frac{\log t}{\log\log t}\right)\right\rfloor
\]
and using the Phragm\'en Lindel\"of principle, one has 
\[
\zeta(\sigma + it) \ll (\log t)^5,\qquad (\sigma \ge 1 - \frac{(\log\log t)^2}{\log t}),
\]
so we may take $f(t) = (\log t)^5$ and $g(t) = (\log\log t)^2/\log t$ in \Cref{mu_to_zero_free}.
\end{proof}

\begin{theorem}[Chudakov zero free region]\label{zfr-chudakov}
One has $\zeta(\sigma + it) \ne 0$ if 
\[
\sigma \ge 1 - \frac{1}{(\log t)^{3/4 + o(1)}}
\]
for $t$ sufficiently large.
\end{theorem}

\begin{theorem}[Korobov-Vinogradov zero free region]\label{zfr-vk}
One has $\zeta(\sigma + it) \ne 0$ if 
\[
\sigma \ge 1 - \frac{A}{(\log t)^{2/3}(\log\log t)^{1/3}}
\]
for an absolute constant $A > 0$ and $t$ sufficiently large.
\end{theorem}
\begin{proof}
Via estimates of Vinogradov's integral, one may obtain an estimate of the form (see e.g. Richert \cite{richert_zur_1967})
\[
\zeta(\sigma + it) \ll t^{B(1 - \sigma)^{3/2}}(\log t)^{O(1)}\qquad (1/2 \le \sigma \le 1)
\]
where $B > 0$ is a constant and $t$ sufficiently large. Take 
\[
g(t) = \left(\frac{\log\log t}{\log t}\right)^{2/3}
\]
so that
\[
\zeta(\sigma + it) \ll f(t) = (\log t)^{O(1)}\qquad (\sigma \ge 1 - g(t)).
\]
The result follows from applying \Cref{mu_to_zero_free}.
\end{proof}

\section{Relation to the error term in the prime number theorem}

\begin{lemma}[Relation to prime number theorem error term]\label{zero_free_to_pnt}
Suppose $\zeta(\sigma + it) \ne 0$ for $\sigma \ge 1 - \eta(t)$ where $\eta(t)$ is a positive and decreasing function. Then
\[
\sum_{n \le x}\Lambda(n) - x \ll x \exp\left(-A \omega(x) \right)\qquad (x \to \infty)
\]
for an absolute constant $A > 0$, where 
\[
\omega(x) := \inf_{t \ge 1}(\eta (t) \log x + \log t).
\]
\end{lemma}
\begin{proof}
See e.g. \cite{ingham_distribution_1990}.
\end{proof}

Applying \Cref{zero_free_to_pnt}, one obtains the error term estimates in the prime number theorem given in Table \ref{zero-free-pnt-table}.   

\begin{table}[ht]
    \def\arraystretch{2.5}
    \centering
    \caption{Error bounds on the prime number theorem. Here $A$ represents an absolute, positive constant, which may be different at each occurrence.}
    \begin{tabular}{|c|c|c|}
    \hline
    Bound on $(\psi(x) - x)/x$ & Associated zero-free region & Reference \\
    \hline
    $\exp(-A(\log x)^{1/2})$ & $\sigma \ge 1 - \dfrac{A}{\log t}$ & \Cref{zfr-classical} \\
    \hline 
    $\exp(-A(\log x \log\log x)^{1/2})$ & $\sigma \ge 1 - \dfrac{A\log\log t}{\log t}$ & \Cref{zfr-littlewood} \\
    \hline 
    $\exp(-A(\log x)^{4/7 + o(1)})$ & $\sigma \ge 1 - \dfrac{A}{(\log t)^{3/4 + o(1)}}$ & \Cref{zfr-chudakov} \\
    \hline 
    $\exp\left(-A\dfrac{(\log x)^{3/5}}{(\log\log x)^{1/5}}\right)$ & $\sigma \ge 1 - \dfrac{A}{(\log t)^{2/3}(\log\log t)^{1/3}}$ & \Cref{zfr-vk} \\
    \hline 
    \end{tabular}
\label{zero-free-pnt-table}
\end{table}

The following type of converse statement is also known.

\begin{theorem}[\cite{turan_new_1984} Theorem 40.1] If for some $0 < \alpha \le 1$ one has
\[
\psi(x) - x \ll x \exp(-A(\log x)^{1/(1 + \alpha)})\qquad (x \to \infty)
\]
then $\zeta(\sigma + it) \ne 0$ for $t$ sufficiently large and
\[
\sigma > 1 - \frac{A}{(\log t)^{\alpha}}.
\]
Here $A$ denotes an absolute positive constant, not necessarily the same at each occurrence. 
\end{theorem}